---
title: "Fitting SDMs"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting_SDMs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  fig.width=6,
  fig.height=6,
  message=FALSE,
  collapse=TRUE
)
```

This vignette shows how to fit SDMs with the **WRAP** package and make some basic plots. The fitting functions in **WRAP** are 

* `gam_sdm()`
* `brt_sdm()`
* `mlp_sdm()`

Type `?gam_sdm`, etc, to learn about the arguments for the function.

# Example

```{r message=FALSE, warning=FALSE}
library(WRAP)
```

## Simulate data 

There are three simulation functions that can be used. `SimulateWorld_ROMS()` which uses ROMS for the environmental fields, `SimulateWorld_ROMS_TrophicInteractins()` which uses ROMS and trophic interactions, and `SimulateWorld()` which simulates a temperature field.

We will use `SimulateWorld()` for this example. Type `?SimulateWorld` as console to learn more about this function.

Parameters for the function are

* `abund_enviro` can be "lnorm_low" (SB); "lnorm_high" (EW); or "poisson" (JS).
* `PA_shape` can be "logistic" (SB); "logistic_prev","linear" (JS)
* `temp_spatial` can be "simple" (SB); or "matern" (EW)
* `temp_diff` is only for the simulated temperature. It specifies min and max temps at year 1 and year 100 (e.g. temp_diff=c(1,3,5,7) means year 1 varies from 1-3C and year 100 from 5-7C). 
* `dir` is the directory where ROMS data is stored. The ROMS functions are looking for a directory called `gfdl`. The default assumption is that this is in your working directory in a folder called `Rasters_2d_monthly`.

We will set the parameters so the following for our simulated abundances and presence/absence grid.

```{r}
abund_enviro <- "lnorm_low"
PA_shape <- "logistic"
temp_spatial <- "matern"
temp_diff <- c(1,4,3,7) 
```

Simulate species abundances from the simulated temperature fields.

```{r simworld, results='hide', message=FALSE}
sim1 <- SimulateWorld(temp_diff = temp_diff,  temp_spatial = temp_spatial, 
                      PA_shape = PA_shape, abund_enviro = abund_enviro) 
```

Printing the output will show you basic information about the simulation. This information is in `sim1$meta`.

```{r print_sim1}
sim1
```

### Save data

Since simulations take a long time to run, you may want to create a directory in your working directory for the simulations and save them there. Then you can reload the simulations instead of re-running.

Create a directory for the simulations:

```{r, eval=FALSE}
if(!dir.exists(file.path(here::here(), 'Sim1')))
  dir.create(file.path(here::here(), 'Sim1'))
```

Save the simulation.

```{r, eval=FALSE}
saveRDS(sim1, file=file.path(here::here(), "Sim1", "Sim1.rds"))
```

## Plot the simulated data

Create dataframe with historical/forecast data. The data frame with the abundances and presence/absence is in `grid` in the simulation object returned by the function.

```{r}
dat <- sim1$grid
dat_hist <- dat[dat$year<=2020,]
dat_fcast <- dat[dat$year>2020,]
```

Make some plots to explore the data. The simulated environmental covariate in `sim1` is `temp`.

```{r, fig.height=6, fig.width=6}
#All Years
par(mfrow=c(2,2))
plot(aggregate(suitability~year,dat,FUN="mean"),type="l", lwd=2, ylab="Suitability",col="dark grey")
lines(aggregate(suitability~year,dat[dat$year<=2020,],FUN="mean"),col="blue")
plot(aggregate(pres~year,dat,FUN="mean"),type="l", lwd=2,ylab="Presence",col="dark grey")
lines(aggregate(pres~year,dat[dat$year<=2020,],FUN="mean"),col="blue")
plot(aggregate(abundance~year,dat,FUN="sum"),type="l",  lwd=2,ylab="Abundance", col="dark grey")
lines(aggregate(abundance~year,dat[dat$year<=2020,],FUN="sum"),col="blue")
plot(aggregate(temp~year,dat,FUN="mean"),type="l",ylab="Temperature", col="dark grey")
```

## Fit SDMs to the simulated grid

We will fit GAM, BRT and MLP SDM models to the simulated grids. The fit function will fit the correct model based on the value of `abund_enviro`. If you want to restrict the smoothing, you can pass in `k` to the GAM function. The default fitting period is years prior to 2021 (`start.forecast.year=2021`) and will use that.

```{r fitmodels, results='hide', message=FALSE, fig.show='hide'}
gam.p <- gam_sdm(sim1, covariates="temp", response="pres")
gam.a <- gam_sdm(sim1, covariates="temp", response="abundance")

brt.p <- brt_sdm(sim1, covariates="temp", response="pres")
brt.a <- brt_sdm(sim1, covariates="temp", response="abundance")

mlp.p <- mlp_sdm(sim1, covariates="temp", response="pres")
mlp.a <- mlp_sdm(sim1, covariates="temp", response="abundance")
```

## Create Predictions

Now plot the comparison of the models for temperaturs 0 to 7C.

First we need to set up the temperatures used for prediction.

```{r}
new_dat <- data.frame(temp=seq(0,7,length=100))
```

Then we will save our predictions from each model. We can do the predictions using our fitted models or we can use the OM object `sim` that was output from the simulation.

Here is how to set up the simulation from the fitted SDM.
```{r pred_fit1, results='hide', message=FALSE, fig.show='hide'}
pred.gam.p <- predict(gam.p, newdata=new_dat, type="response")
pred.gam.a <- predict(gam.a, newdata=new_dat, type="response")

pred.brt.p <- predict(brt.p, newdata=new_dat, type="response",
     n.trees=brt.p$gbm.call$best.trees)
pred.brt.a <- predict(brt.a, newdata=new_dat, type="response",
     n.trees=brt.a$gbm.call$best.trees)

dat_norm <- new_dat
dat_norm$temp <- BBmisc::normalize(dat_norm$temp)
pred.mlp.p <- predict(mlp.p, dat_norm)
pred.mlp.a <- predict(mlp.a, dat_norm)
```

Alternatively here is how to use the fitted SDM to create predictions. You can (optionally) pass in a fitted model. If you do not, the predict function with fit a SDM based on the `model` argument. We will pass in the fitted SDMs to speed up the predictions.  This predict function for a OM object returns a data frame with pred.p, pred.a, pred and CIs added. A column pred.type indicates the type of prediction: hindcast, forecast, or newdata.

```{r alt_pred, results='hide', message=FALSE, fig.show='hide'}
pred.gam <- predict(sim1, p.sdm=gam.p, a.sdm=gam.a, newdata=new_dat)
pred.brt <- predict(sim1, p.sdm=brt.p, a.sdm=brt.a, newdata=new_dat)
pred.mlp <- predict(sim1, p.sdm=mlp.p, a.sdm=mlp.a, newdata=new_dat)
```

## Plot Predictions

```{r}
par(mfrow=c(2,2), mar=c(3,4,4,2))
ylim2 <- 1.05*max(pred.gam$pred, pred.brt$pred, pred.mlp$pred)

#actual TPC
xx <- seq(0, 7, length=100)
#yy must match response function in SimulateWorld function (sim1$meta$response.curve$temp)
yy <- dnorm(xx, mean=4, sd=1)  
plot(xx, yy, type="l", lty=2, main="Actual TPC", col="red", xlim=c(0,8), ylab="suitability", xlab="Temp")
xlim <- round(100*(max(new_dat$temp)/7))
lines(xx[1:xlim], yy[1:xlim], lwd=2)

#gam
plot(new_dat$temp, pred.gam$pred, type="l",
     main="lnorm_low GAM", xlim=c(0,8), col="red", lty=2, ylim=c(0,ylim2), ylab="Abundance", xlab="Temp")
lines(new_dat$temp, pred.gam$pred, lwd=2)

#brt
plot(new_dat$temp, pred.brt$pred, type="l",
     main="lnorm_low BRT", xlim=c(0,8), col="red", lty=2, ylim=c(0,ylim2), ylab="Abundance", xlab="Temp")
lines(new_dat$temp, pred.brt$pred, lwd=2)

#MLP
plot(new_dat$temp, pred.mlp$pred, type="l",
     main="lnorm_low MLP", xlim=c(0,8), col="red", lty=2, ylim=c(0,ylim2), ylab="Abundance", xlab="Temp")
lines(new_dat$temp, pred.mlp$pred, lwd=2)

par(mfrow=c(1,1))
```

## Make Hindcast and Forecast predictions

If `newdata` is left off, `predict(sim)` where `sim` is the output from one of the simulation functions (operating model) will return a data frame with a hindcast for years before start.forecast.year and a forecast after that. See `?predict.OM` for other arguments that this function takes.

You can pass in a previously fitted SDM to speed up the function as we did before. But you can leave these off and `predict` will fit the SDM based on the value in `sim$meta$abund_enviro`. Specify the SDM model with argument `model`.

```{r pred_om, results='hide', fig.show='hide'}
pred.gam2 <- predict(sim1, model="gam")
pred.brt2 <- predict(sim1, model="brt")
pred.mlp2 <- predict(sim1, model="mlp")
```

Assemble the predictions into one data frame.
```{r}
nr <- nrow(pred.gam2)
pred.all <- rbind(pred.gam2, pred.brt2, pred.mlp2)
pred.all$model <- c(rep("gam", nr), rep("brt", nr), rep("mlp", nr))
```

## Compare abundance predictons

First we create a data frame of the abundance predictions and true abundance.

```{r}
abund_enviro <- sim1$meta$abund_enviro

# sum up the abundance predictions across cells by year
# model col is the model
abund <- aggregate(pred~year+model, pred.all, FUN="sum")
colnames(abund) <- c("year", "model", "abundance")

# Add true abundance
x <- sim1$grid
if (abund_enviro == "poisson") x$abundance <- round(abundance)
tmp <- aggregate(abundance~year, x, FUN="sum")
tmp$model <- "true"
tmp <- tmp[,c("year", "model", "abundance")]

abund <- rbind(abund, tmp)
```

Next we plot

```{r}
p <- ggplot(abund, aes(x=year, y=abundance, color=model)) + 
  geom_line() +
  ggtitle("Comparison of lnorm_low models") +
  geom_vline(xintercept=2020) +
  annotate("text", x=2020, y=max(abund$abundance), label="  forecast", hjust=0) +
  annotate("text", x=2020, y=max(abund$abundance), label="hindcast  ", hjust=1)
```


## Calculate and plot centre of gravity

```{r}
# First compute the true cog
library(dplyr)
cog_lat <- x %>% group_by(year) %>% 
  summarize(cog=weighted.mean(x=lat, w=abundance))
cog_lat <- cbind(model="true", cog_lat, stringsAsFactors = FALSE)

# Now add the model cogs
tmp <- pred.all %>% group_by(model, year) %>% 
  summarize(cog=weighted.mean(x=lat, w=pred))

# dplyr uses tibbles and they return a matrix when you use rbind(). ug.
# so use 
cog_lat <- dplyr::bind_rows(cog_lat, tmp)
```

`cog_lat` is now a data frame with the center of gravity and a model column for the model (or true).

```{r}
p <- ggplot(cog_lat, aes(x=year, y=cog, color=model)) + 
  geom_line() +
  ggtitle("Comparison of lnorm_low models") +
  ylab("Centre of Gravity (deg lat)") +
  geom_vline(xintercept=2020) +
  annotate("text", x=2020, y=max(cog_lat$cog), label="  forecast", hjust=0) +
  annotate("text", x=2020, y=max(cog_lat$cog), label="hindcast  ", hjust=1)

p
```

## Plot Surface Predictions

These are point predictions for individual cells.  We will show this for a particular year, in this case 2021.

```{r surf_pred, fig.height=6, fig.width=6}
#Future
Y = 2021
x <- subset(sim1$grid, year==Y)
#Truth
p1 <- ggplot(x, aes(lon, lat))+
  geom_tile(aes(fill=abundance)) +
  theme_classic() +
  ggtitle("Truth")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  scale_fill_viridis()

#Gam
x <- subset(pred.all, year==Y & model=="gam")
p2 <- ggplot(x, aes(lon,lat))+
  geom_tile(aes(fill=pred)) +
  theme_classic() +
  ggtitle("GAM")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  scale_fill_viridis()

#BRT
x <- subset(pred.all, year==Y & model=="brt")
p3 <- ggplot(x, aes(lon,lat))+
  geom_tile(aes(fill=pred)) +
  theme_classic() +
  ggtitle("BRT")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  scale_fill_viridis()

#Gam
x <- subset(pred.all, year==Y & model=="mlp")
p4 <- ggplot(x, aes(lon,lat))+
  geom_tile(aes(fill=pred)) +
  theme_classic() +
  ggtitle("MLP")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  scale_fill_viridis()

gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
```
