---
title: "Run Operating Model"
author: "E Holmes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulate_Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  fig.width=6,
  fig.height=6,
  message=FALSE,
  collapse=TRUE
)
```

The simulation functions are

* `SimulateWorld()` Use `?SimulateWorld` to read documentation
* `SimulateWorld_ROMS()`
* `SimulateWorld_ROMS_TrophicInteraction()`

## Run simulation

### Use `SimulateWorld()`

First load the **WRAP** library.

```{r message=FALSE, warning=FALSE}
library(WRAP)
```

This function simulates presence and abundance using a simulated temperature field.  Parameters for the function are 

* `abund_enviro` can be "lnorm_low" (SB); "lnorm_high" (EW); or "poisson" (JS).
* `PA_shape` can be "logistic" (SB); "logistic_prev","linear" (JS)
* `temp_spatial` can be "simple" (SB); or "matern" (EW)
* `temp_diff` is only for the simulated temperature. It specifies min and max temps at year 1 and year 100 (e.g. temp_diff=c(1,3,5,7) means year 1 varies from 1-3C and year 100 from 5-7C). 

The defaults are  `temp_diff = c(1, 3, 5, 7)`, `temp_spatial = "simple"`, `PA_shape = "logistic"`, and `abund_enviro = "lnorm_low"`.  You can also set number of years, grid size, and start year.

```{r run_sim_1, message=FALSE, results='hide'}
sim <- SimulateWorld()
```

`sim` (the output from a simulation function) is class "OM" and is a list. It has the simulated grid in `grid` and information about the simulation in `meta`. Typing the name of the output will print basic information about the simulation.

```{r print_sim}
sim
```

### Saving the simulation

### Save data

Since simulations take a long time to run, you may want to create a directory in your working directory for the simulations and save them there. Then you can reload the simulations instead of re-running.

Create a directory for the simulations:

```{r, eval=FALSE}
if(!dir.exists(file.path(here::here(), 'Sim1')))
  dir.create(file.path(here::here(), 'Sim1'))
```

Save the simulation.

```{r, eval=FALSE}
saveRDS(sim, file=file.path(here::here(), "Sim1", "Sim.rds"))
```

## Plot

Plot time-series of total catch in observed years

```{r plot_obs}
data <- sim$grid
plot(aggregate(abundance~year, data[data$year<=2010,], FUN="sum"),type="l",ylab="Abundance")
```

plot time-series of total catch in forecast years

```{r plot_fore}
data <- sim$grid
plot(aggregate(abundance~year,data[data$year>2010,], FUN="sum"),type="l", ylab="Abundance")
```

## Plot the abundance

```{r plot_grid_2001, fig.height=4, fig.width=4, warning=FALSE}
library(raster)
df <- subset(sim$grid, year==2001)
df <- df[, c("lon", "lat", "abundance")]
dfr <- raster::rasterFromXYZ(df)      
plot(dfr)
```

Show 4 years and remove the box and legend.
 
```{r plot_grid_4, fig.height=6, fig.width=6, warning=FALSE}
library(raster)
par(mfrow=c(2,2), mar=c(2,2,2,2))
for(yr in c(2001, 2025, 2050, 2100)){
df <- subset(sim$grid, year==yr)
df <- df[, c("lon", "lat", "abundance")]
dfr <- raster::rasterFromXYZ(df)      
plot(dfr, axes=FALSE, box=FALSE, legend=FALSE)
title(yr)
}
```


