---
title: "Plotting OMs and SDMs"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make_Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  fig.width=6,
  fig.height=6,
  message=FALSE,
  collapse=TRUE
)
```

This vignette shows how to make plots of OMs (simulated rasters of environmental variables and species abundances) and SDMs (fitted species distribution models) using the  **WRAP** package. 

**WRAP** includes the following plotting functions 

* `plot()` Plot yearly averages of the variables in an OM or POM
* `plot_cog()` Center of gravity for each year for OMs, SDMs and POMs
* `plot_abund()` Total abundance for each year for OMs, SDMs and POMs
* `plot_grid()` Grid of abundance for one or a range of years for OMs, SDMs and POMs
* `plot_pres()` Average presence for each year for OMs, SDMs and POMs

This vignette will show the **WRAP** functions and also show you how to make the plots from your own code.

# Example

```{r message=FALSE, warning=FALSE}
library(WRAP)
```

## Simulate data 

For theses examples, we'll use an OM from `SimulateWorld_ROMS()` which uses ROMS for the environmental fields.  Set the parameters to the following for our simulated abundances and presence/absence grid. `PA_shape`	specifies how enviro suitability determines species presence-absence. It takes values of "logistic", "logistic_prev", or "linear". `abund_enviro`	specifies abundance relationship when presence=1. It can be "lnorm_low", "lnorm_high", or "poisson".

```{r}
abund_enviro <- "lnorm_low"
PA_shape <- "logistic"
```

Simulate species abundances:

```{r simworld, results='hide', message=FALSE}
sim1 <- SimulateWorld_ROMS(PA_shape = PA_shape, 
                      abund_enviro = abund_enviro,
                      roms.years = 2000:2040) 
```

Printing the output will show you basic information about the simulation. This information is in `sim1$meta`.

```{r print_sim1}
sim1
```


## Plot the yearly averages

Get the yearly means of the variables in an OM model.

```{r}
plot(sim1, start.forecast.year=2020)
```

Here is raw code to make the same plot.

```{r}
library(dplyr) #for the %>% pipe
year_mean <- sim1$grid %>% aggregate(.~year, ., FUN="mean") %>% 
  reshape2::melt(., id.vars=c("year"),
                  measure.vars=c("pres", "suitability", "sst", "abundance"),
                  variable.name="variable",
                  value.name="value")
year_mean$date <- as.Date(paste(year_mean$year, "01-01", sep="-"))
```

Plot. Blue is the training data and grey is the testing (forecast) data.

```{r}
ggplot(year_mean, aes(x=date, y=value)) + geom_line(col="grey") +
  geom_line(aes(x=date, y=value), col="blue", subset(year_mean, year<=2020)) +
  facet_wrap(~variable, scales = "free_y") +
  xlab("year")
```

Here is an example without `ggplot()`.

```{r, fig.height=6, fig.width=6}
#All Years
dat <- sim1$grid
par(mfrow=c(2,2))
plot(aggregate(suitability~year,dat,FUN="mean"),type="l", lwd=2, ylab="Suitability",col="dark grey")
lines(aggregate(suitability~year,dat[dat$year<=2020,],FUN="mean"),col="blue")
plot(aggregate(pres~year,dat,FUN="mean"),type="l", lwd=2,ylab="Presence",col="dark grey")
lines(aggregate(pres~year,dat[dat$year<=2020,],FUN="mean"),col="blue")
plot(aggregate(abundance~year,dat,FUN="sum"),type="l",  lwd=2,ylab="Abundance", col="dark grey")
lines(aggregate(abundance~year,dat[dat$year<=2020,],FUN="sum"),col="blue")
plot(aggregate(sst~year,dat,FUN="mean"),type="l",ylab="Temperature", col="dark grey")
```

## Fit SDMs to the simulated grid

We will fit GAM, BRT and MLP SDM models to the simulated grids. The default fitting period is years prior to 2021 (`start.forecast.year=2021`).

```{r fitmodels, results='hide', message=FALSE, fig.show='hide'}
gam.fit <- gam_sdm(sim1, covariates="sst")
brt.fit <- brt_sdm(sim1, covariates="sst")
mlp.fit <- mlp_sdm(sim1, covariates="sst")
```

## Compare abundance predictons

The function `plot_abund()` will plot the total abundances of the true data (from simulation) and the fitted models. The functions works on OMs (from the simulation functions), SDMs (from the fitting functions), and POMs (from the prediction function).

```{r}
plot_abund(sim1, gam.fit, mlp.fit, brt.fit)
# You can also pass in predictions, which would be faster
```

Here is how to compute these abundance plots without `plot_abund`. First we create a data frame of the abundance predictions and true abundance.

```{r pred, results='hide', message=FALSE, fig.show='hide'}
pred.gam <- predict(sim1, sdm=gam.fit)
pred.brt <- predict(sim1, sdm=brt.fit)
pred.mlp <- predict(sim1, sdm=mlp.fit)

# Assemble into a data frame
nr <- nrow(pred.gam)
pred.all <- rbind(pred.gam, pred.brt, pred.mlp)
pred.all$model <- c(rep("gam", nr), rep("brt", nr), rep("mlp", nr))
```

```{r}
# sum up the abundance predictions across cells by year
# model col is the model
abund <- aggregate(pred~year+model, pred.all, FUN="sum")
colnames(abund) <- c("year", "model", "abundance")

# Add true abundance
x <- sim1$grid
if (sim1$meta$abund_enviro == "poisson") x$abundance <- round(x$abundance)
tmp <- aggregate(abundance~year, x, FUN="sum")
tmp$model <- "true"
tmp <- tmp[,c("year", "model", "abundance")]

abund <- rbind(abund, tmp)
```

Next we plot

```{r}
p <- ggplot(abund, aes(x=year, y=abundance, color=model)) + 
  geom_line() +
  ggtitle("Comparison of models") +
  geom_vline(xintercept=2020) +
  annotate("text", x=2020, y=max(abund$abundance), label="  forecast", hjust=0) +
  annotate("text", x=2020, y=max(abund$abundance), label="hindcast  ", hjust=1)
p
```


## Calculate and plot centre of gravity

The function `plot_cog()` will plot the center of gravity of OMs (from simulations) and fitted models. If you pass in a
fitted object (SDM), `plot_cog()` will run `predict()` on that object or you can pass in
a predict object (POM) returned by `predict()` if you already have that. The latter will
be faster since the plotting doesn't need to rerun `predict()`. You can pass in any combination of OMs, SDMs, and POMs.

```{r}
plot_cog(sim1, pred.gam, pred.mlp, pred.brt)
# You can also pass in SDMs
# plot_cog(sim1, gam.fit, mlp.fit, brt.fit)
```

Here is how to compute these COG plots without `plot_cog`. First we create a data frame of the COG predictions and true abundance.


```{r}
# First compute the true cog
library(dplyr)
cog_lat <- x %>% group_by(year) %>% 
  summarize(cog=weighted.mean(x=lat, w=abundance))
cog_lat <- cbind(model="true", cog_lat, stringsAsFactors = FALSE)

# Now add the model cogs
tmp <- pred.all %>% group_by(model, year) %>% 
  summarize(cog=weighted.mean(x=lat, w=pred))

# dplyr uses tibbles and they return a matrix when you use rbind().
# so use 
cog_lat <- dplyr::bind_rows(cog_lat, tmp)
```

`cog_lat` is now a data frame with the center of gravity and a model column for the model (or true).

```{r}
p <- ggplot(cog_lat, aes(x=year, y=cog, color=model)) + 
  geom_line() +
  ggtitle("Comparison of lnorm_low models") +
  ylab("Centre of Gravity (deg lat)") +
  geom_vline(xintercept=2020) +
  annotate("text", x=2020, y=max(cog_lat$cog), label="  forecast", hjust=0) +
  annotate("text", x=2020, y=max(cog_lat$cog), label="hindcast  ", hjust=1)

p
```

## Plot average presence

The function `plot_pres()` will plot average presence by year. This function is similar to `plot_abund()`.

```{r}
plot_pres(sim1, pred.gam, pred.mlp, pred.brt)
# You can also pass in SDMs
# plot_cog(sim1, gam.fit, mlp.fit, brt.fit)
```

## Plot Surface Predictions

These are point predictions for individual cells.  Since there are only 400 points per year for this simulation, it will look sparse if we only plot one year. We will plot 20 years, 2021 to 2040.  Note `plot_grid()` works with any combination of OM, SDM and POM objects, but if SDMs are included then the first argument must be the OM for which predictions will be made from the SDM.


```{r}
plot_grid(sim1, year=2021:2040, pred.gam, pred.mlp, pred.brt)
# You can also pass in SDMs but the predictions will be computed (slow)
# plot_grid(sim1, year=2021:2040, gam.fit, mlp.fit, brt.fit)
```

Here is how you could make the plot directly.

```{r surf_pred, fig.height=6, fig.width=6}
#Future
uscale <- 9.5
Y <- 2021:2040
x <- subset(sim1$grid, year%in%Y)
#Truth
p1 <- ggplot(x, aes(x=lon, y=lat))+
  geom_tile(aes(fill=abundance)) +
  theme_classic() +
  ggtitle("Truth")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  viridis::scale_fill_viridis(limits=c(0, uscale))

#Gam
x <- subset(pred.all, year%in%Y & model=="gam")
p2 <- ggplot(x, aes(lon,lat))+
  geom_tile(aes(fill=pred)) +
  theme_classic() +
  ggtitle("GAM")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  viridis::scale_fill_viridis(limits=c(0, uscale))

#BRT
x <- subset(pred.all, year%in%Y & model=="brt")
p3 <- ggplot(x, aes(lon,lat))+
  geom_tile(aes(fill=pred)) +
  theme_classic() +
  ggtitle("BRT")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  viridis::scale_fill_viridis(limits=c(0, uscale))

#Gam
x <- subset(pred.all, year%in%Y & model=="mlp")
p4 <- ggplot(x, aes(lon,lat))+
  geom_tile(aes(fill=pred)) +
  theme_classic() +
  ggtitle("MLP")+
  labs(y="Latitude") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous( expand = c(0, 0)) +
  theme(legend.title=element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  viridis::scale_fill_viridis(limits=c(0, uscale))

gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
```
